---
title: 'Homework 2: Data 605'
author: "Umer Farooq"
date: "2024-10-26"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1:

### Bayesian:

**Problem Statement:**

A new credit scoring system has been developed to predict the likelihood of loan defaults. The system has a 90% sensitivity, meaning that it correctly identifies 90% of those who will default on their loans. It also has a 95% specificity, meaning that it correctly identifies 95% of those who will not default. The default rate among borrowers is 2%. Given these prevalence, sensitivity, and specificity estimates, what is the probability that a borrower flagged by the system as likely to default will actually default? If the average loss per defaulted loan is $200,000 and the cost to run the credit scoring test on each borrower is $500, what is the total first-year cost for evaluating 10,000 borrowers?

**Solution:**

A credit scoring system has been developed with the following metrics:
- \textbf{Sensitivity} (\(P(\text{Positive Test} \mid \text{Default})\)) = 90\%
- \textbf{Specificity} (\(P(\text{Negative Test} \mid \text{No Default})\)) = 95\%
- \textbf{Prevalence of Default} (\(P(\text{Default})\)) = 2\%

We need to calculate:
1. The probability that a borrower flagged as likely to default will actually default.
2. The total cost of evaluating 10,000 borrowers, where:
   - The average loss per defaulted loan is \$200,000.
   - The cost to run the test for each borrower is \$500.

\section*{Solution}

\subsection*{1. Probability of Default Given a Positive Test}

Using Bayes' Theorem, we have:

\[
P(\text{Default} \mid \text{Positive Test}) = \frac{P(\text{Positive Test} \mid \text{Default}) \cdot P(\text{Default})}{P(\text{Positive Test})}
\]

Where:
\[
P(\text{Positive Test}) = P(\text{Positive Test} \mid \text{Default}) \cdot P(\text{Default}) + P(\text{Positive Test} \mid \text{No Default}) \cdot P(\text{No Default})
\]

Given:
\[
P(\text{Positive Test} \mid \text{Default}) = 0.90, \quad P(\text{Default}) = 0.02
\]
\[
P(\text{Positive Test} \mid \text{No Default}) = 1 - \text{Specificity} = 1 - 0.95 = 0.05
\]
\[
P(\text{No Default}) = 1 - P(\text{Default}) = 1 - 0.02 = 0.98
\]

Substitute these values:

\[
P(\text{Positive Test}) = (0.90 \times 0.02) + (0.05 \times 0.98) = 0.018 + 0.049 = 0.067
\]

Thus:

\[
P(\text{Default} \mid \text{Positive Test}) = \frac{0.90 \times 0.02}{0.067} \approx 0.2687 \text{ or } 26.87\%
\]

\subsection*{2. Total First-Year Cost for Evaluating 10,000 Borrowers}

Given:
- Average loss per defaulted loan = \$200,000
- Cost of test per borrower = \$500
- Number of borrowers = 10,000

First, calculate the total test cost:

\[
\text{Total Test Cost} = 10,000 \times 500 = 5,000,000
\]

Then, calculate the expected loss cost:

\[
\text{Expected Defaults in Flagged Positives} = P(\text{Default} \mid \text{Positive Test}) \times 10,000 \times 0.02
\]
\[
= 0.2687 \times 10,000 \times 0.02 = 53.74
\]

\[
\text{Expected Loss Cost} = 53.74 \times 200,000 \approx 10,747,462.69
\]

So, the total cost for the first year is:

\[
\text{Total First-Year Cost} = \text{Total Test Cost} + \text{Expected Loss Cost} = 5,000,000 + 10,747,462.69 = 15,747,462.69
\]

\end{document}



```{r}
# Given values
sensitivity <- 0.90  # P(Test Positive | Default)
specificity <- 0.95  # P(Test Negative | No Default)
prevalence <- 0.02   # P(Default)
non_default_prevalence <- 1 - prevalence  # P(No Default)

# Calculate P(Test Positive)
p_test_positive <- (sensitivity * prevalence) + ((1 - specificity) * non_default_prevalence)

# Calculate P(Default | Test Positive) using Bayes' Theorem
ppv <- (sensitivity * prevalence) / p_test_positive
ppv

```

```{r}
# Cost values
loss_per_default <- 200000  # Average loss per defaulted loan
test_cost_per_borrower <- 500  # Cost to run the credit scoring test on each borrower
num_borrowers <- 10000  # Number of borrowers

# Expected defaults in flagged positives (PPV * flagged positives)
expected_defaults_in_flagged <- ppv * num_borrowers * prevalence

# Total cost: Test cost for all borrowers + Expected losses from defaults
total_test_cost <- num_borrowers * test_cost_per_borrower
expected_loss_cost <- expected_defaults_in_flagged * loss_per_default
total_first_year_cost <- total_test_cost + expected_loss_cost

total_first_year_cost

```

***

### Binomial:

**Problem Statement:**

The probability that a stock will pay a dividend in any given quarter is 0.7. What is the probability that the stock pays dividends exactly 6 times in 8 quarters? What is the probability that it pays dividends 6 or more times? What is the probability that it pays dividends fewer than 6 times? What is the expected number of dividend payments over 8 quarters? What is the standard deviation?

**Solution:**

The probability that a stock will pay a dividend in any given quarter is 0.7. Over 8 quarters, we need to calculate:
1. The probability that the stock pays dividends exactly 6 times.
2. The probability that it pays dividends 6 or more times.
3. The probability that it pays dividends fewer than 6 times.
4. The expected number of dividend payments over 8 quarters.
5. The standard deviation of dividend payments.

The dividend payments follow a **Binomial distribution** where:
\[
n = 8 \quad \text{and} \quad p = 0.7
\]

\section*{Solution}

\subsection*{1. Probability of Exactly 6 Dividends in 8 Quarters}

Using the Binomial probability formula:
\[
P(X = k) = \binom{n}{k} p^k (1 - p)^{n - k}
\]

Given \( n = 8 \), \( k = 6 \), and \( p = 0.7 \), we find \( P(X = 6) \):

\[
P(X = 6) = \binom{8}{6} (0.7)^6 (0.3)^2
\]
\[
= \frac{8!}{6!(8-6)!} \times (0.7)^6 \times (0.3)^2
\]
\[
\approx 0.2965
\]

Thus, the probability that the stock pays dividends exactly 6 times in 8 quarters is approximately 0.2965.

\subsection*{2. Probability of 6 or More Dividends in 8 Quarters}

To find \( P(X \geq 6) \), we sum the probabilities for 6, 7, and 8 dividends:
\[
P(X \geq 6) = P(X = 6) + P(X = 7) + P(X = 8)
\]

Using a binomial cumulative distribution:
\[
P(X \geq 6) \approx 0.5518
\]

So, the probability that it pays dividends 6 or more times is approximately 0.5518.

\subsection*{3. Probability of Fewer Than 6 Dividends}

To find \( P(X < 6) \), we sum the probabilities for 0 through 5 dividends:
\[
P(X < 6) = P(X = 0) + P(X = 1) + \dots + P(X = 5)
\]

Using the cumulative distribution function:
\[
P(X < 6) \approx 0.4482
\]

Thus, the probability that it pays dividends fewer than 6 times is approximately 0.4482.

\subsection*{4. Expected Number of Dividend Payments}

The expected number of successes \( E(X) \) in a Binomial distribution is given by:
\[
E(X) = n \cdot p
\]

Substitute \( n = 8 \) and \( p = 0.7 \):
\[
E(X) = 8 \times 0.7 = 5.6
\]

So, the expected number of dividend payments over 8 quarters is 5.6.

\subsection*{5. Standard Deviation of Dividend Payments}

The standard deviation \( \sigma \) for a Binomial distribution is given by:
\[
\sigma = \sqrt{n \cdot p \cdot (1 - p)}
\]

Substitute \( n = 8 \) and \( p = 0.7 \):
\[
\sigma = \sqrt{8 \times 0.7 \times 0.3} \approx 1.2961
\]

Thus, the standard deviation of dividend payments over 8 quarters is approximately 1.2961.

\end{document}

R-Code:

```{r}
# Given parameters
n <- 8         # Number of trials (quarters)
p <- 0.7       # Probability of a dividend payment per quarter

# 1. Probability of exactly 6 dividends in 8 quarters
k <- 6
prob_6_dividends <- dbinom(k, n, p)
cat("Probability of exactly 6 dividends in 8 quarters:", round(prob_6_dividends, 4), "\n")

# 2. Probability of 6 or more dividends in 8 quarters
# P(X >= 6) = P(X = 6) + P(X = 7) + P(X = 8)
prob_6_or_more_dividends <- pbinom(5, n, p, lower.tail = FALSE)
cat("Probability of 6 or more dividends in 8 quarters:", round(prob_6_or_more_dividends, 4), "\n")

# 3. Probability of fewer than 6 dividends in 8 quarters
# P(X < 6) = P(X <= 5)
prob_fewer_than_6_dividends <- pbinom(5, n, p)
cat("Probability of fewer than 6 dividends in 8 quarters:", round(prob_fewer_than_6_dividends, 4), "\n")

# 4. Expected number of dividend payments
expected_dividends <- n * p
cat("Expected number of dividend payments over 8 quarters:", expected_dividends, "\n")

# 5. Standard deviation of dividend payments
std_deviation_dividends <- sqrt(n * p * (1 - p))
cat("Standard deviation of dividend payments:", round(std_deviation_dividends, 4), "\n")

```

***

### Poisson: 

**Problem Statement: **

A financial analyst notices that there are an average of 12 trading days each month when a certain stock’s price increases by more than 2%. What is the probability that exactly 4 such days occur in a given month? What is the probability that more than 12 such days occur in a given month? How many such days would you expect in a 6-month period? What is the standard deviation of the number of such days? If an investment strategy requires at least 70 days of such price increases in a year for profitability, what is the percent utilization and what are your recommendations?

**Solution: **

A financial analyst observes that there are, on average, 12 trading days each month when a certain stock’s price increases by more than 2%. We need to calculate:

1. The probability that exactly 4 such days occur in a given month.
2. The probability that more than 12 such days occur in a given month.
3. The expected number of such days in a 6-month period.
4. The standard deviation of the number of such days.
5. If an investment strategy requires at least 70 days of such price increases in a year for profitability, calculate the percent utilization and provide recommendations.

We assume that the number of days follows a **Poisson distribution** with a mean \( \lambda = 12 \) (days per month).

\section*{Solution}

\subsection*{1. Probability of Exactly 4 Days in a Month}

Using the Poisson probability formula:
\[
P(X = k) = \frac{\lambda^k e^{-\lambda}}{k!}
\]

Given \( \lambda = 12 \) and \( k = 4 \), we find \( P(X = 4) \):
\[
P(X = 4) = \frac{12^4 e^{-12}}{4!}
\]
\[
\approx 0.0027
\]

So, the probability of exactly 4 days in a month with a price increase of over 2% is approximately 0.27%.

\subsection*{2. Probability of More Than 12 Days in a Month}

To find \( P(X > 12) \), we use:
\[
P(X > 12) = 1 - P(X \leq 12)
\]

Using cumulative probability for the Poisson distribution:
\[
P(X > 12) \approx 0.384
\]

Therefore, the probability of more than 12 such days is approximately 38.4%.

\subsection*{3. Expected Number of Days in a 6-Month Period}

The expected value for a Poisson distribution over 6 months is:
\[
E(X) = 6 \cdot \lambda = 6 \cdot 12 = 72
\]

Thus, we expect approximately 72 days over a 6-month period.

\subsection*{4. Standard Deviation of the Number of Days}

The standard deviation \( \sigma \) for a Poisson distribution is given by:
\[
\sigma = \sqrt{\lambda}
\]

For a 6-month period:
\[
\sigma = \sqrt{6 \cdot 12} = \sqrt{72} \approx 8.49
\]

So, the standard deviation over 6 months is approximately 8.49 days.

\subsection*{5. Percent Utilization and Recommendation}

If an investment strategy requires at least 70 days of such price increases in a year, the expected number of days is:
\[
E(X) = 12 \cdot 12 = 144 \quad \text{(days per year)}
\]

The percent utilization is:
\[
\text{Utilization} = \frac{70}{144} \times 100 \approx 48.6\%
\]

**Recommendation**: Given that the percent utilization is approximately 48.6%, the strategy is likely feasible if the target is 70 days, as the expected value is well above this threshold. However, due to variability, further analysis on monthly variances or a contingency plan is advised.

\end{document}

R-Code:

```{r}
# Given parameters
lambda <- 12   # Average number of days per month with price increase > 2%

# 1. Probability of exactly 4 days in a month
k <- 4
prob_4_days <- dpois(k, lambda)
cat("Probability of exactly 4 days in a month:", round(prob_4_days, 4), "\n")

# 2. Probability of more than 12 days in a month
prob_more_than_12 <- ppois(12, lambda, lower.tail = FALSE)
cat("Probability of more than 12 days in a month:", round(prob_more_than_12, 4), "\n")

# 3. Expected number of days in a 6-month period
expected_6_months <- 6 * lambda
cat("Expected number of days in a 6-month period:", expected_6_months, "\n")

# 4. Standard deviation of the number of days in a 6-month period
std_dev_6_months <- sqrt(expected_6_months)
cat("Standard deviation of days over 6 months:", round(std_dev_6_months, 4), "\n")

# 5. Percent utilization and recommendation
# Expected number of days in a year
expected_year <- 12 * lambda
target_days <- 70
percent_utilization <- (target_days / expected_year) * 100
cat("Percent utilization for 70 target days:", round(percent_utilization, 2), "%\n")

```

***

### Hypergeometric

**Problem Statement:**

A hedge fund has a portfolio of 25 stocks, with 15 categorized as high-risk and 10 as low-risk. The fund manager randomly selects 7 stocks to closely monitor. If the manager selected 5 high-risk stocks and 2 low-risk stocks, what is the probability of selecting exactly 5 high-risk stocks if the selection was random? How many high-risk and low-risk stocks would you expect to be selected?

**Solution:**

A hedge fund has a portfolio of 25 stocks, with 15 categorized as high-risk and 10 as low-risk. The fund manager randomly selects 7 stocks to monitor. We are asked to determine:

1. The probability that exactly 5 of the selected stocks are high-risk.
2. The expected number of high-risk and low-risk stocks in the sample.

This scenario follows a **Hypergeometric distribution**, where:
- \( N = 25 \) (total number of stocks),
- \( K = 15 \) (number of high-risk stocks),
- \( n = 7 \) (number of stocks selected),
- \( k = 5 \) (number of high-risk stocks chosen in the sample).

\section*{Solution}

\subsection*{1. Probability of Selecting Exactly 5 High-Risk Stocks}

The probability of selecting exactly 5 high-risk stocks is given by the Hypergeometric formula:
\[
P(X = k) = \frac{\binom{K}{k} \cdot \binom{N - K}{n - k}}{\binom{N}{n}}
\]

Substitute \( N = 25 \), \( K = 15 \), \( n = 7 \), and \( k = 5 \):
\[
P(X = 5) = \frac{\binom{15}{5} \cdot \binom{10}{2}}{\binom{25}{7}}
\]

Calculating each component:
\[
\binom{15}{5} = \frac{15!}{5!(15 - 5)!} = 3003, \quad \binom{10}{2} = 45, \quad \binom{25}{7} = 480700
\]

Substitute these values:
\[
P(X = 5) = \frac{3003 \times 45}{480700} \approx 0.2812
\]

Thus, the probability of selecting exactly 5 high-risk stocks is approximately 0.2812.

\subsection*{2. Expected Number of High-Risk and Low-Risk Stocks}

For a Hypergeometric distribution, the expected number of high-risk stocks in the sample is:
\[
E(\text{high-risk}) = n \cdot \frac{K}{N} = 7 \cdot \frac{15}{25} = 4.2
\]

Similarly, the expected number of low-risk stocks is:
\[
E(\text{low-risk}) = n \cdot \frac{N - K}{N} = 7 \cdot \frac{10}{25} = 2.8
\]

Therefore, we expect to select approximately 4.2 high-risk and 2.8 low-risk stocks.

\end{document}

R-Code:


```{r}
# Given parameters
N <- 25          # Total number of stocks
K <- 15          # Number of high-risk stocks
n <- 7           # Number of stocks selected
k <- 5           # Number of high-risk stocks chosen

# 1. Probability of selecting exactly 5 high-risk stocks
prob_5_high_risk <- dhyper(k, K, N - K, n)
cat("Probability of exactly 5 high-risk stocks:", round(prob_5_high_risk, 4), "\n")

# 2. Expected number of high-risk and low-risk stocks
expected_high_risk <- n * (K / N)
expected_low_risk <- n * ((N - K) / N)
cat("Expected number of high-risk stocks:", round(expected_high_risk, 2), "\n")
cat("Expected number of low-risk stocks:", round(expected_low_risk, 2), "\n")

```

***

### Geometric:

**Problem Statement:**

The probability that a bond defaults in any given year is 0.5%. A portfolio manager holds this bond for 10 years. What is the probability that the bond will default during this period? What is the probability that it will default in the next 15 years? What is the expected number of years before the bond defaults? If the bond has already survived 10 years, what is the probability that it will default in the next 2 years?

**Solution:**

The probability that a bond defaults in any given year is 0.5%. A portfolio manager holds this bond for a specified period. We are asked to determine:

1. The probability that the bond will default within 10 years.
2. The probability that the bond will default within 15 years.
3. The expected number of years before the bond defaults.
4. The probability that the bond defaults within 2 years, given that it has already survived for 10 years.

This scenario follows a **Geometric distribution**, where:
- \( p = 0.005 \) (probability of default each year).

\section*{Solution}

\subsection*{1. Probability of Default within 10 Years}

For a Geometric distribution, the probability that the bond defaults within \( n \) years is given by the cumulative distribution function:
\[
P(X \leq n) = 1 - (1 - p)^n
\]

For \( p = 0.005 \) and \( n = 10 \):
\[
P(X \leq 10) = 1 - (1 - 0.005)^{10}
\]
\[
= 1 - (0.995)^{10} \approx 0.0488
\]

Thus, the probability that the bond will default within 10 years is approximately 4.88%.

\subsection*{2. Probability of Default within 15 Years}

Similarly, for \( n = 15 \):
\[
P(X \leq 15) = 1 - (1 - 0.005)^{15}
\]
\[
= 1 - (0.995)^{15} \approx 0.0729
\]

The probability of default within 15 years is approximately 7.29%.

\subsection*{3. Expected Number of Years Before Default}

The expected number of years before default for a Geometric distribution is given by:
\[
E(X) = \frac{1}{p}
\]

Substituting \( p = 0.005 \):
\[
E(X) = \frac{1}{0.005} = 200
\]

Thus, the expected number of years before default is 200 years.

\subsection*{4. Probability of Default within 2 Years Given Survival for 10 Years}

Given that the bond has already survived for 10 years, the probability that it defaults within the next 2 years is:
\[
P(X \leq 2) = 1 - (1 - p)^2
\]

For \( p = 0.005 \):
\[
P(X \leq 2) = 1 - (0.995)^2 \approx 0.009975
\]

Therefore, the probability that the bond defaults in the next 2 years, given it has survived 10 years, is approximately 0.9975%.

\end{document}

R-Code:

```{r}
# Given parameters
p <- 0.005      # Probability of default in a given year

# 1. Probability of default within 10 years
n1 <- 10
prob_within_10_years <- 1 - (1 - p)^n1
cat("Probability of default within 10 years:", round(prob_within_10_years, 4), "\n")

# 2. Probability of default within 15 years
n2 <- 15
prob_within_15_years <- 1 - (1 - p)^n2
cat("Probability of default within 15 years:", round(prob_within_15_years, 4), "\n")

# 3. Expected number of years before default
expected_years_before_default <- 1 / p
cat("Expected number of years before default:", expected_years_before_default, "\n")

# 4. Probability of default within 2 years given survival for 10 years
n3 <- 2
prob_within_2_years_given_10_years <- 1 - (1 - p)^n3
cat("Probability of default within 2 years after surviving 10 years:", round(prob_within_2_years_given_10_years, 4), "\n")

```

***

###Poisson: 

**Problem Statement:**

A high-frequency trading algorithm experiences a system failure about once every 1500 trading hours. What is the probability that the algorithm will experience more than two failures in 1500 hours? What is the expected number of failures?

**Solution:**

A high-frequency trading algorithm experiences a system failure once every 1500 trading hours on average. We need to calculate:

1. The probability that the algorithm will experience more than two failures in 1500 hours.
2. The expected number of failures over 1500 hours.

Since this is a rare event over a fixed interval, we can model the number of failures with a **Poisson distribution**. The average rate (\( \lambda \)) for 1500 hours is 1 failure.

\section*{Solution}

\subsection*{1. Expected Number of Failures}

The expected number of failures in 1500 hours, given that failures occur on average once every 1500 hours, is:
\[
E(X) = \lambda = 1
\]

So, we expect 1 failure over 1500 hours.

\subsection*{2. Probability of More Than 2 Failures in 1500 Hours}

For a Poisson distribution, the probability of observing more than \( k \) events is calculated as:
\[
P(X > k) = 1 - P(X \leq k)
\]

In this case, we need the probability of more than 2 failures (\( k = 2 \)).

First, we calculate \( P(X \leq 2) \), the cumulative probability for 0, 1, and 2 failures:
\[
P(X \leq 2) = P(X = 0) + P(X = 1) + P(X = 2)
\]

Each term in the Poisson probability mass function is given by:
\[
P(X = k) = \frac{\lambda^k e^{-\lambda}}{k!}
\]

For \( \lambda = 1 \):
\[
P(X = 0) = \frac{1^0 e^{-1}}{0!} = e^{-1} \approx 0.3679
\]
\[
P(X = 1) = \frac{1^1 e^{-1}}{1!} = e^{-1} \approx 0.3679
\]
\[
P(X = 2) = \frac{1^2 e^{-1}}{2!} = \frac{e^{-1}}{2} \approx 0.1839
\]

Now, summing these probabilities:
\[
P(X \leq 2) = 0.3679 + 0.3679 + 0.1839 = 0.9197
\]

Then,
\[
P(X > 2) = 1 - P(X \leq 2) = 1 - 0.9197 = 0.0803
\]

Thus, the probability of experiencing more than two failures in 1500 hours is approximately 8.03%.

\end{document}

R-Code:

```{r}
# Given parameter
lambda <- 1  # Mean number of failures in 1500 hours

# 1. Expected number of failures
expected_failures <- lambda
cat("Expected number of failures in 1500 hours:", expected_failures, "\n")

# 2. Probability of more than 2 failures
prob_more_than_2 <- 1 - ppois(2, lambda)
cat("Probability of more than 2 failures in 1500 hours:", round(prob_more_than_2, 4), "\n")

```

### Uniform Distribution: 

**Problem Statement:**

An investor is trying to time the market and is monitoring a stock that they believe has an equal chance of reaching a target price between 20 and 60 days. What is the probability that the stock will reach the target price in more than 40 days? If it hasn’t reached the target price by day 40, what is the probability that it will reach it in the next 10 days? What is the expected time for the stock to reach the target price?

**Solution:**

An investor monitors a stock expected to reach a target price between 20 and 60 days with equal probability for any day in this range. We are asked to calculate:

1. The probability that the stock will reach the target price in more than 40 days.
2. Given it hasn’t reached the target price by day 40, the probability it will reach it in the next 10 days.
3. The expected time for the stock to reach the target price.

Since the stock reaching the target price is uniformly distributed, we can use the properties of the **Uniform distribution** on the interval \([a, b]\).

\section*{Solution}

\subsection*{1. Probability of Reaching the Target Price in More Than 40 Days}

For a continuous uniform distribution over \([a, b]\), the probability that the time \( X \) is greater than a certain day \( c \) is:
\[
P(X > c) = \frac{b - c}{b - a}
\]

Substituting \( c = 40 \), \( a = 20 \), and \( b = 60 \):
\[
P(X > 40) = \frac{60 - 40}{60 - 20} = \frac{20}{40} = 0.5
\]

So, the probability that the stock will reach the target price in more than 40 days is 0.5 (or 50%).

\subsection*{2. Probability of Reaching the Target Price Between Days 40 and 50, Given It Hasn't Reached by Day 40}

Given that the target hasn’t been reached by day 40, we are interested in the conditional probability that it reaches the target in the next 10 days (between days 40 and 50). This is calculated as:
\[
P(40 < X \leq 50 \mid X > 40) = \frac{P(40 < X \leq 50)}{P(X > 40)}
\]

We already know \( P(X > 40) = 0.5 \). Now we calculate \( P(40 < X \leq 50) \) as:
\[
P(40 < X \leq 50) = \frac{50 - 40}{60 - 20} = \frac{10}{40} = 0.25
\]

Thus,
\[
P(40 < X \leq 50 \mid X > 40) = \frac{0.25}{0.5} = 0.5
\]

So, given it hasn’t reached by day 40, the probability of reaching the target in the next 10 days is also 0.5 (or 50%).

\subsection*{3. Expected Time to Reach the Target Price}

The expected value \( E(X) \) for a uniform distribution on \([a, b]\) is given by:
\[
E(X) = \frac{a + b}{2}
\]

Substituting \( a = 20 \) and \( b = 60 \):
\[
E(X) = \frac{20 + 60}{2} = \frac{80}{2} = 40
\]

So, the expected time for the stock to reach the target price is 40 days.

\end{document}

R-Code:

```{r}
# Given parameters
a <- 20  # Minimum time in days
b <- 60  # Maximum time in days
c <- 40  # Given day for probability calculations

# 1. Probability that the stock reaches the target price in more than 40 days
prob_more_than_40 <- (b - c) / (b - a)
cat("Probability of reaching the target price in more than 40 days:", prob_more_than_40, "\n")

# 2. Probability that the stock reaches the target in the next 10 days given it hasn’t by day 40
prob_40_to_50 <- (50 - 40) / (b - a)    # Probability of reaching between day 40 and 50
cond_prob_40_to_50 <- prob_40_to_50 / prob_more_than_40
cat("Conditional probability of reaching target between days 40 and 50:", cond_prob_40_to_50, "\n")

# 3. Expected time to reach the target price
expected_time <- (a + b) / 2
cat("Expected time to reach the target price:", expected_time, "\n")

```

### Exponential Distribution: 

**Problem Statement:**

A financial model estimates that the lifetime of a successful start-up before it either goes public or fails follows an exponential distribution with an expected value of 8 years. What is the expected time until the start-up either goes public or fails? What is the standard deviation? What is the probability that the start-up will go public or fail after 6 years? Given that the start-up has survived for 6 years, what is the probability that it will go public or fail in the next 2 years?


**Solution:**

A financial model estimates that the lifetime of a successful start-up before it either goes public or fails follows an Exponential distribution with an expected lifetime of 8 years. We need to calculate:

1. The expected time until the start-up either goes public or fails.
2. The standard deviation of the lifetime.
3. The probability that the start-up will go public or fail after 6 years.
4. Given that the start-up has survived for 6 years, the probability it will go public or fail in the next 2 years.

\section*{Solution}

\subsection*{1. Expected Time Until the Start-up Goes Public or Fails}

For an Exponential distribution, the expected value \( E(X) \) is given as:
\[
E(X) = \frac{1}{\lambda}
\]

Since the expected lifetime is already provided as 8 years, we have:
\[
E(X) = 8
\]

\subsection*{2. Standard Deviation of the Lifetime}

The standard deviation of an Exponential distribution is also \( \frac{1}{\lambda} \), which is equal to the expected value. Therefore:
\[
\text{Standard Deviation} = 8
\]

\subsection*{3. Probability that the Start-up Will Go Public or Fail After 6 Years}

For an Exponential distribution, the probability that the lifetime \( X \) exceeds a specific time \( t \) is:
\[
P(X > t) = e^{-\lambda t}
\]

Substituting \( \lambda = 0.125 \) and \( t = 6 \):
\[
P(X > 6) = e^{-0.125 \times 6} = e^{-0.75} \approx 0.4724
\]

So, the probability that the start-up will go public or fail after 6 years is approximately 0.4724.

\subsection*{4. Probability of Going Public or Failing in the Next 2 Years Given Survival for 6 Years}

This is a conditional probability calculation. For an Exponential distribution, the memoryless property allows us to calculate it as:
\[
P(X > 6 \text{ and } X \leq 8 \mid X > 6) = P(X \leq 2) = 1 - e^{-\lambda \times 2}
\]

Substituting \( \lambda = 0.125 \):
\[
P(X \leq 2) = 1 - e^{-0.125 \times 2} = 1 - e^{-0.25} \approx 0.2212
\]

Thus, given that the start-up has already survived 6 years, there is approximately a 22.12% probability that it will go public or fail in the next 2 years.

\end{document}

R-Code:

```{r}
# Given parameters
expected_lifetime <- 8   # Expected lifetime in years
lambda <- 1 / expected_lifetime  # Rate parameter

# 1. Expected time until the start-up goes public or fails
expected_time <- expected_lifetime
cat("Expected time until going public or failure:", expected_time, "years\n")

# 2. Standard deviation of the lifetime
std_dev <- 1 / lambda
cat("Standard deviation of the lifetime:", std_dev, "years\n")

# 3. Probability that the start-up will go public or fail after 6 years
t1 <- 6
prob_after_6_years <- exp(-lambda * t1)
cat("Probability of going public or failing after 6 years:", round(prob_after_6_years, 4), "\n")

# 4. Probability of going public or failing in the next 2 years given survival for 6 years
t2 <- 2
prob_next_2_years_given_6 <- 1 - exp(-lambda * t2)
cat("Probability of going public or failing in the next 2 years given survival for 6 years:", round(prob_next_2_years_given_6, 4), "\n")

```

